### 2017-05-22 sheets/basis-3/sheet-basis-3.tex
Some solutions

\begin{exercise}
  Define categories $\cat{A}$ and $\cat{B}$ such that either $\cat{A} \iso \cat{B}$ in $\CAT$ or $\cat{A} \eqv \cat{B}$, but not both.
  Do not use the example given during the lecture.
\end{exercise}

@@

\begin{exercise}
  Let $A$ and $B$ be sets, and denote by $B^A$ the set of functions from $A$
  to $B$.  Write down:
  \begin{itemize}
  \item[(a)] a canonical function $A \to B^{(B^A)}$,
  \item[(b)] a canonical function $A \times B^{A} \to B$.
  \end{itemize}
\end{exercise}

@@

\begin{exercise}
        Let $\xymatrix@1{\cat{A}\rtwocell^F_G{\alpha} &\cat{B}}$ be a natural transformation.
        \begin{itemize}
                \item[(a)] Suppose $\alpha$ is a natural isomorphism.
                Show that $\alpha_A\from F(A) \to G(A)$ is an isomorphism for all $A \in \cat{A}$.
                \item[(b)] Now assume that $\alpha_A\from F(A) \to G(A)$ is an isomorphism for all
                $A \in \cat{A}$. Prove that $\alpha$ is a natural isomorphism.
        \end{itemize}
\end{exercise}

@@

Fix a field $K$.  Let $\Mt$ be the category whose objects are the natural
numbers and with
\[ \Mt(m, n) = \{ n \times m \text{ matrices over } K \}. \]
Prove that $\Mt$ is equivalent to $\FDVect_K$, the category of
finite-dimensional vector spaces over $K$.
Does your equivalence involve a \emph{canonical} functor from $\Mt$ to $\FDVect_K$, or from $\FDVect_K$ to $\Mt$?
\end{exercise}

>>>>

\begin{exercise}
  Define categories $\cat{A}$ and $\cat{B}$ such that either $\cat{A} \iso \cat{B}$ in $\CAT$ or $\cat{A} \eqv \cat{B}$, but not both.
  Do not use the example given during the lecture.
\end{exercise}
\begin{answer}
    \begin{minipage}{0.5\textwidth}
        $\cat{A}$:
        \[
            \begin{tikzcd}
                A \arrow{d}{f}\\
                B
            \end{tikzcd}
        \]
    \end{minipage}
    \begin{minipage}{0.5\textwidth}
        $\cat{B}$:
        \[
            \begin{tikzcd}
                A' \arrow{dr}{f'} \arrow[bend left]{rr}{g} &    & A'' \arrow{dl}{f''} \arrow{ll}{g^{-1}} \\
                                                           & B' &
            \end{tikzcd}
        \]
    \end{minipage}
    The identity morphisms have been left out for simplicity.
    First of all note that there is no bijection between the objects of $\cat{A}$ and $\cat{B}$, so $\cat{A} \not\iso \cat{B}$.\\
    We now show $\cat{A} \eqv \cat{B}$.
    Therefore we define a functor $F$:
    \begin{align*}
        F(A) := A'\\
        F(B) := B'\\
        F(f) := f'
    \end{align*}
    In addition $F$ maps the identity morhisms to their respective identity morphisms.
    It is easy to check that $F$ is full and faithful.
    Furthermore $F$ is surjective on the objects, because $A'$ and $A''$ are isomorph in $\cat{B}$.
    Thus $\cat{A} \eqv \cat{B}$.
\end{answer}

@@

\begin{exercise}
  Let $A$ and $B$ be sets, and denote by $B^A$ the set of functions from $A$
  to $B$.  Write down:
  \begin{itemize}
  \item[(a)] a canonical function $A \to B^{(B^A)}$,
  \item[(b)] a canonical function $A \times B^{A} \to B$.
  \end{itemize}
\end{exercise}
\begin{answer}
    \begin{itemize}
        \item[(a)] $\lambda\, a\, \lambda\,f.\, f\,a$
        \item[(b)] $\lambda\, (a, f).\, f\,a$
    \end{itemize}
\end{answer}

@@

\begin{exercise}
        Let $\xymatrix@1{\cat{A}\rtwocell^F_G{\alpha} &\cat{B}}$ be a natural transformation.
        \begin{itemize}
            \item[(a)] Suppose $\alpha$ is a natural isomorphism.
                Show that $\alpha_A\from F(A) \to G(A)$ is an isomorphism for all $A \in \cat{A}$.
            \item[(b)] Now assume that $\alpha_A\from F(A) \to G(A)$ is an isomorphism for all
                $A \in \cat{A}$. Prove that $\alpha$ is a natural isomorphism.
        \end{itemize}
\end{exercise}
\begin{answer}
    \begin{itemize}
        \item[(a)]
            By definition of natural isomorphism we have a natural transformation $\beta: G \Rightarrow F$, such that
            \[\beta \circ \alpha = 1_{F} \text{ and } \alpha \circ \beta = 1_{G}. \]
            We show that $\forall A \in \cat{A}.\, \beta_A \circ \alpha_A = 1_{F(A)} \text{ and } \alpha_A \circ \beta_A = 1_{G(A)}$.
            Suppose $A \in \cat{A}$.
            We know by definition of the identity natural transformation that $(\beta \circ \alpha)_A = 1_{F(A)}$.
            The claim $\beta_A \circ \alpha_A = 1_{F(A)}$ follows by definition of composition of natural transformations.
            $\alpha_A \circ \beta_A = 1_{G(A)}$ is shown analogously.
        \item[(ab)]
            We know that for all $A \in \cat{A}$ there is a $\beta_A$ such that
            \[\beta_A \circ \alpha_A = 1_{F(A)} \text{ and } \alpha_A \circ \beta_A = 1_{G(A)}.\]
            First we prove that $\beta$ is a natural transformation.
            Suppose $f\from A \to A'$.
            We know that $\alpha$ is a natural transformation:
            \begin{align*}
                & G(f) \circ \alpha_A && = \alpha_{A'} \circ F(f)\\
                \Rightarrow\, & G(f) \circ \alpha_A \circ \beta_A && = \alpha_{A'} \circ F(f) \circ \beta_A\\
                \Rightarrow\, & \beta_{A'} \circ G(f) && = \beta_{A'} \circ \alpha_{A'} \circ F(f) \circ \beta_A\\
                \Rightarrow\, & \beta_{A'} \circ G(f) && = F(f) \circ \beta_A
            \end{align*}
            By similar reasoning to (a) we get that $\beta$ is the inverse of $\alpha$.
            Therefore $\alpha$ is a natural isomorpism.
    \end{itemize}
\end{answer}

@@

Fix a field $K$.  Let $\Mt$ be the category whose objects are the natural
numbers and with
\[ \Mt(m, n) = \{ n \times m \text{ matrices over } K \}. \]
Prove that $\Mt$ is equivalent to $\FDVect_K$, the category of
finite-dimensional vector spaces over $K$.
Does your equivalence involve a \emph{canonical} functor from $\Mt$ to $\FDVect_K$, or from $\FDVect_K$ to $\Mt$?
\end{exercise}
\begin{answer}
    Define a functor $F: \Mt \to \FDVect_K$:
    \begin{align*}
        F(n) & := K^n && \forall n \in \NN\\
        F(A) & := \lambda\, x. \, A \cdot x && \forall n \times m \text{ matrices } A
    \end{align*}
    Note that $F$ is indeed a functor, since matrix multiplication is associative.
    \begin{description}
        \item[$F$ is faithful:]
            Let $A, B$ be two $n \times m$ matrices over $K$ such that $F(A) = F(B)$.
            By using the standard basis in $K^m$ we get, that each column of $A$ and $B$ are the same.
            Thus $A = B$.
        \item[$F$ is full:]
            Let $f: K^m \to K^n$ be a linear function.
            We define the $n \times m$ matrix $A$ columnwise.
            The $j$-th column of $A$ is $f(e_j)$, where $e_j$ is the $j$-th standard basis vector.
            By this definition $F(A)$ and $f$ conincide on the standard basis.
            Linear algebra tells us that in this case $F(A)$ and $f$ coincide on all vectors.
            So $A$ is mapped to $f$ under $F$.
        \item[$F$ is surjective on the objects:]
            Let $V$ be a finite-dimensional $K$-vectorspace.
            Then $V$ has some dimension $n \in \NN$.
            Thus $V \iso K^n = F(n)$.
    \end{description}
\end{answer}
